### 2020-02-22 ISLR/08_Tree-Based-Methods.rmd
wrapped up trees
wrapped up trees

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(tree, quietly = TRUE, warn.conflicts = FALSE)
library(randomForest, quietly = TRUE, warn.conflicts = FALSE)
library(gbm, quietly = TRUE, warn.conflicts = FALSE)

@@

```{r}
yhat.bag <- predict(bag.boston, newdata = boston[-train])
plot(yhat.bag, boston.test)
abline(0, 1)
mean((yhat.bag-boston.test)^2)
```

```{r}
bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = boston[-train,])
mean((yhat.bag - boston.test)^2)
```

@@

ntrees <- seq(0, 500, 10)

results <- numeric(length(ntrees))

for(i in 1:length(ntrees))
{
   tree.count <- ntrees[i]

   rf <- randomForest(medv ~ ., data = boston.train, mtry = 6, n.trees = tree.count)

@@

   results[i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
}

rf.results <- data.table(trees = ntrees, errors = results)[, lowest := .I == lowest.error]

ggplot(rf.results, aes(trees, errors, fill = lowest)) +
   geom_line()

@@

@@

>>>>

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(tree, quietly = TRUE, warn.conflicts = FALSE)
library(randomForest, quietly = TRUE, warn.conflicts = FALSE)
library(gbm, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

@@

```{r}
boston.test <- boston[!train]
yhat.bag <- predict(bag.boston, newdata = boston.test)
plot(yhat.bag, boston.test$medv)
abline(0, 1)
mean((yhat.bag-boston.test$medv)^2)
```

```{r}
bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = boston[-train,])
mean((yhat.bag - boston.test$medv)^2)
```

@@

ntrees <- seq(0, 500, 10)
n.features <- ncol(boston[, !"medv"])

mtry <- round(c(n.features, n.features/2, sqrt(n.features)))

results <- matrix(vector(mode = "numeric", length(ntrees) * length(mtry)), nrow = 3)

for(i in 1:length(ntrees))
{
   for(j in 1:length(mtry))
   {
      tree.count <- ntrees[i]

      rf <- randomForest(medv ~ ., data = boston.train, mtry = mtry[j], n.trees = tree.count)

@@

      results[j, i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
   }
}

df_results <- as.data.table(results)
colnames(df_results) <- paste0(ntrees)

df_results <- melt(cbind(mtry, df_results), id.vars = "mtry")
df_results$mtry <- as.factor(df_results$mtry)

ggplot(df_results, aes(variable, value, group = mtry)) +
   geom_line(aes(col = mtry))

@@

### Carseats

@@

### OJ

This problem involves the **OJ** data set which is part of the ISLR package.

a.) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.

```{r}
oj <- as.data.table(ISLR::OJ)
n <- nrow(oj)

index <- sample(1:n, n *.7, replace = F)

oj.train <- oj[index]
oj.test <- oj[!index]

head(train)
```

b.) Fit a tree to the training data, which **Purchases** as the first response and the other variables as predictors. Use the **summary** function to produce summary statistics about the tree, and describe the results obtained.

oj.tree <- tree(Purchase ~., data = oj.train)

plot(oj.tree)
text(oj.tree)

summary(oj.tree)

What is the training error rate? _0.7469_

How many terminal nodes? _8_

c.) Type in the name of the tree object in order to get a detailed text output. Pick one of the terminal nodes, and interpret the information displayed.

```{r}
summary(oj.tree)
```

d.) Create a plot of the tree.

```{r}
plot(oj.tree)
text(oj.tree)
```

e.) Predict the response on the test data, and produce a confusion matrix comparing the test labels to the predicted test labels.

```{r}
pred <- predict(oj.tree, newdata = oj.test, type = "class")

confusionMatrix(oj.test$Purchase, pred)
```

f.) Apply the cv.tree function to the training data set in order to determine the optimal tree size.

```{r}
oj.cv <- cv.tree(oj.tree)
```

g.) Produce a plot with the tree size and cross-validated error rate.

```{r}
plot(oj.cv$size, oj.cv$dev, type ='b', col =  "cornflowerblue") # 5
```

h.) What tree size corresponds to the lowest cv error rate?

**5**

i.) Produce a pruned tree corresponding to the optimal tree size obtained using cross-validation.

```{r}
oj.pruned <- prune.tree(oj.tree, k = 5)
```

j.) Compare the training error rates between the pruned and unpruned trees.

summary(oj.tree)
summary(oj.pruned)
```

Same.

k.) Compare the test error rates.

```{r}
pruned.pred <- predict(oj.pruned, newdata = oj.test, type = "class")

confusionMatrix(oj.test$Purchase, pred)

confusionMatrix(oj.test$Purchase, pruned.pred)
```

### Hitters

We now use boosting to predict **Salary** in the hitters data set.

a.) Remove the observations for whom the salary information is unknown, and then log-transform the salaries.

```{r}
hitters <- as.data.table(ISLR::Hitters)
hitters <- hitters[complete.cases(hitters)]
hitters$Salary <- log(hitters$Salary)

b.) Create a training set consisting of the first 200 observations and a test set containing the rest.

```{r}
n <- nrow(hitters)

index <- 1:200

hitters.train <- hitters[index]
hitters.test <- hitters[!index]
```

c.) Perform boosting on the training set with 1,000 trees for a range of values of the shrinkage parameter $\lambda$. Produce a plot with the different shrinkage values and the training MSE.

```{r}
shrinkage <- seq(0.001, 0.5, by = 0.0025)
iters <- length(shrinkage)

results <- vector(mode = "numeric", length = iters)

for(i in 1:iters)
{
   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")
   results[i] <- mean(gbm$train.error^2)
}

df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]

ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
   geom_bar(stat = "identity")
```

d.) Produce a plot with different shrinkage methods and the corresponding test mse.

```{r}
shrinkage <- seq(0.001, 0.5, by = 0.0025)
iters <- length(shrinkage)

results <- vector(mode = "numeric", length = iters)

for(i in 1:iters)
{
   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")

   pred <- predict(gbm, n.trees = 1000, newdata = hitters.test, type = "response")

   results[i] <- mean((hitters.test$Salary - pred)^2)
}

df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]

ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
   geom_bar(stat = "identity")

tree.mse <- results[lowest.error]
```

e.) Compare the test MSE of boosting to the test MSE that results from applying two of the regression approaches seen in chapters 3 & 6.

train.mat <- model.matrix(Salary ~., data = hitters.train)
test.mat <- model.matrix(Salary ~., data = hitters.test)

cv.ridge <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 0)

ridge.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 1, lambda = cv.ridge$lambda.min)
ridge.pred <- predict(ridge.fit, newx = test.mat, type = "response")
ridge.error <- mean((hitters.test$Salary - ridge.pred)^2)

cv.lasso <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 1)

lasso.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 0, lambda = cv.lasso$lambda.min)
lasso.pred <- predict(lasso.fit, newx = test.mat, type = "response")
lasso.error <- mean((hitters.test$Salary - lasso.pred)^2)

results <- data.table(type = c("Tree", "Ridge", "Lasso"), error = c(tree.mse, ridge.error, lasso.error))

ggplot(results, aes(type, error, fill = type)) +
   geom_bar(stat = "identity")
```

f.) Which variables appear to be the most important predictors in the boosted model?

best.tree <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[lowest.error], distribution = "gaussian")

vip::vip(best.tree)
```

### Caravan

This question uses the caravan data set.

a.) Create a training set consisting of the first 1,000 observations, and a test set consisting of the remaining observations.

```{r}
caravan <- as.data.table(ISLR::Caravan)
caravan$Purchase <- ifelse(caravan$Purchase == "Yes", 1, 0)
index <- 1:1000

caravan.train <- caravan[index]
caravan.test <- caravan[-index]

head(caravan.train)
```

b.) Fit a boosting model to the training set with **Purchases** as the response and the other variables as predictors. Use 1,000 trees and a shrinkage value of .01.

```{r}
boost.fit <- gbm(Purchase ~ ., data = caravan.train, n.trees = 1000, shrinkage = 0.1, distribution = "bernoulli")
```

What predictors appear to be most important?

```{r}
vip::vip(boost.fit)
```

c.) Use the boosting model to predict the response on the test data. Predict that a person will make a purchase if the estimated probablity of purchase is greater than 20%.

```{r}
probs <- exp(predict.gbm(boost.fit, newdata = caravan.test, type = "response", n.trees = 1000)) - 1

pred <- ifelse(probs > .2, 1, 0)

confusionMatrix(as.factor(caravan.test$Purchase), as.factor(pred))
```
